Microbial interactions as drivers of a nitrifying ecosystem in a chemostat

### 2020-01-21 tex/Article_Interactions.tex
added several corrections

The control $u(t)$ is intended to drive the system to be near a desired output $z(t)$, which in this context are the measurements of the concentrations of OTU. The term $\lambda \Vert(u -\vec{1}) \Vert_R$, was added for two reasons:
\begin{itemize}
\item First because the interest is testing the idea that interactions could be driving the system. Therefore adding a penalization in the objective function for each control to remain near $1$ can be seen as an attempt to explain data without any interaction.
\item Second, to force a regularized control. Otherwise note that $u$ is linear in \eqref{Controlsystem}, therefore if the integral cost does not have a non-linear expression of $u$ the optimal control will be of a bang-bang type with possibly singular arcs. Since the objective is to find a differentiable expression of $I(x)$ the addition of the regularization term is deemed necessary.
\end{itemize} 
The problem of approximating the solution of the system to a desired reference ($z$ in this case) is called the optimal tracking problem. For solving such a problem the approach developed by Cimen \textit{et al.} \cite{Cimen2004, Cimen2008} was adapted to our problem. The method proposed involves the resolution of Approximating Sequences of Ricatti Equations (ASRE).

@@

The change of variable $u_i(t) = v_i(t) + 1$ is used for technical reasons explained in the appendix section. This in turn implies $v_i(t) \in [-1,a-1]$.

@@

The approach was tested against synthetic data. The test consisted in simulating system \eqref{system} with growth functions \eqref{gLV growth1} and \eqref{gLV growth2} (therefore $I(x) = 1 + Ax$) for $n_1 = 10$ and $n_2 = 5$, with a randomly generated matrix $A$ in figure \ref{Interaction Matrix Simulation}. In all figures, asterisks represent the synthetic data, while the continuous lines represent the method's output.

The choice of matrices $Q$ and $R$ are always of discussion in a quadratic regulator. $Q$ represents the covariance matrix of measurements, while the interpretation of $R$ is not clear. In the simulations here presented $Q = I_n$ and $R = \lambda I_n$. $\lambda$ is varied from $10^{-1}$ to $10^{-7}$ in the supplementary material section. The case $\lambda = 10^{-4}$ is shown in here for discussion.

Figures \ref{Total Biomass no noise}, \ref{AOB no noise}, and \ref{NOB no noise} show the total biomass concentration, and the trajectories for the OTU belonging to $G_1$ and $G_2$, respectively. It can be seen that the method approaches well the trajectories. The metabolites concentration represented in figure \ref{Metabolites no noise} are in accordance with the simulated: The method is able to reconstruct the metabolites trajectories from the community measurements. 
\begin{figure}[h]
	\centering	
	\includegraphics[width=0.5\textwidth]{Synthetic_data_no_noise//Interactions_synthetic_data_no_noise}
	\caption{Interaction matrix associated with model \eqref{system} for a simulated consortia of 15 bacterial species where $x_1$-$x_{10}$ are AOB and $x_{11}$-$x_{15}$ are NOB.}
	\label{Interaction Matrix Simulation}
\end{figure}

@@

The interaction function introduced tries to address the issue of prediction from genetic sequencing. The analysis of the gLV model proposed by Dumont \textit{et al.} \cite{Dumont2016} shows a new mechanism by which the CEP may not hold, however we conjecture that the system seems to have only one positive stable equilibrium. From the application section one can see that by correcting the growth rate one is able to recuperate the chemical species dynamics, implying that interactions are driving the system. The former correction varies strongly and when compared to the proof of section one may hypothesise that linear pairwise interaction model such as the gLV model can not capture the complexity of the particular chemostat analysed. This opens the question for a broader class of interaction functions that could fit well complex microbial ecosystems. The experimental study in chemostat systems seems a good compromise between completely open ecosystems (nature) and controlled conditions (laboratory) in order to explore expressions that could account for microbial interactions.

>>>>

The control $u(t)$ is intended to drive the system to be near a desired output $z(t)$, which in this context are the measurements of the concentrations of OTU. The term $\lambda \Vert(u -\vec{1}) \Vert_R$, was added for two reasons:
\begin{itemize}
\item First because the interest is testing the idea that interactions could be driving the system. Therefore adding a penalization in the objective function for each control to remain near $1$ can be seen as an attempt to explain data without any interaction.
\item Second, to force a regularized control. Otherwise note that $u$ is linear in \eqref{Controlsystem}, therefore if the integral cost does not have a non-linear expression of $u$ the optimal control will be of a bang-bang type with possibly singular arcs \cite{harmand2019optimal}. Since the objective is to find a differentiable expression of $I(x)$ the addition of the regularization term is deemed necessary.
\end{itemize} 
The problem of approximating the solution of the system to a desired reference ($z$ in this case) is called the optimal tracking problem. For solving such a problem the approach developed by Cimen \textit{et al.} \cite{Cimen2004, Cimen2008} was adapted to our problem. The method proposed involves the resolution of Approximating Sequences of Ricatti Equations (ASRE).

@@

The change of variable $u_i(t) = v_i(t) + 1$ is used for technical reasons explained in the supplementary material section. This in turn implies $v_i(t) \in [-1,\bar{u}-1]$.

@@

The approach was tested against synthetic data. The test consisted in simulating system \eqref{system} with growth functions \eqref{gLV growth1} and \eqref{gLV growth2}, therefore the interaction function $I(x) = 1 + Ax$), $f_i(s) = \dfrac{s_1}{K_i +s_1}$ for $i \in G_1$, and $f_i(s) = \dfrac{s_2}{K_i +s_2}$ for $i \in G_2$, for $n_1 = 10$ and $n_2 = 5$, with a randomly generated matrix $A$ in figure \ref{Interaction Matrix Simulation}. Parameters of the simulation can be seen in table . A simulation of the synthetic data is presented in figure \ref{fig:nonoise200121aobplot} and figure \ref{fig:nonoise200121nobplot} depending on operating parameters $D$ and $s_{in}$.

\begin{table}[ht]
	\centering
	\begin{tabular}{|l|l|l|l|}
		\hline
		Modified Parameters  & $\mu_i\,[1/day]$ & $K_i\,[mg/L]$ & $k_i =\frac{1}{y_i} \, [gr/gr]$ \\ \hline
		$x_i \in G_1$ &0.77  &0.07  & 3.94 \\ \hline
		$x_i\in G_2$ & 0.621 & 0.0147 & 16.12 \\ \hline
	\end{tabular}
	\caption{Kinetic parameters of model \eqref{system} modified from \cite{Dumont2016}.}
	\label{Parameters_synthetic_data}
\end{table}

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.45 \linewidth}
	\includegraphics[width=\linewidth]{Images/proof_of_concept/no_noise_200121_AOB_plot}
	\caption{}
	\label{fig:nonoise200121aobplot}
	\end{subfigure}
	\begin{subfigure}{0.45 \linewidth}
	\centering
	\includegraphics[width=\linewidth]{Images/proof_of_concept/no_noise_200121_NOB_plot}
	\caption{}
	\label{fig:nonoise200121nobplot}
	\end{subfigure}
\caption{}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\linewidth]{Images/proof_of_concept/no_noise_200121_metabolites}
	\caption{}
	\label{fig:nonoise200121metabolites}
\end{figure}

\begin{figure}[h]
	\centering	
	\begin{subfigure}{0.45 \linewidth}
		\includegraphics[width=\linewidth]{Synthetic_data_no_noise//Interactions_synthetic_data_no_noise}
		\caption{Interaction matrix associated with model \eqref{system} for a simulated consortia of 15 bacterial species where $x_1$-$x_{10}$ are AOB and $x_{11}$-$x_{15}$ are NOB.}
		\label{Interaction Matrix Simulation}
	\end{subfigure}
\end{figure}

In all figures, asterisks represent the synthetic data, while the continuous lines represent the method's output.

For applying the optimal tracking approach, the choice of matrices $Q$ and $R$ are always of discussion in a quadratic regulator. $Q$ represents the covariance matrix of measurements, while $R$ is subject to interpretations depending on the context: in the case here it does not have a definite meaning, however its behaviour is studied when $\rho(R) \rightarrow 0$, with $\rho(R)$ being the spectral radius of matrix $R$. In the simulations here presented $Q = I_n$ and $R = \lambda I_n$. $\lambda$ is varied from $10^{-1}$ to $10^{-7}$ in the supplementary material section. The case $\lambda = 10^{-4}$ is shown in here for discussion.

Figures \ref{Total Biomass no noise}, \ref{AOB no noise}, and \ref{NOB no noise} show the total biomass concentration, and the trajectories for the OTU belonging to $G_1$ and $G_2$, respectively. It can be seen that the method approaches well the trajectories. The metabolites concentration represented in figure \ref{Metabolites no noise} are in accordance with the simulated: The method is able to reconstruct the metabolites trajectories from the community measurements.

@@

The interaction function introduced tries to address the issue of prediction from genetic sequencing. The analysis of the gLV model proposed by Dumont \textit{et al.} \cite{Dumont2016} shows a new mechanism by which the CEP may not hold, however we conjecture that the system seems to have only one positive stable equilibrium. From the application section one can see that by correcting the growth rate one is able to recuperate the substrates dynamics, implying that interactions are driving the system. The former correction varies strongly and when compared to the proof of section one may hypothesise that linear pairwise interaction model such as the gLV model can not capture the complexity of the particular chemostat analysed. This opens the question for a broader class of interaction functions that could fit well complex microbial ecosystems. The experimental study in chemostat systems seems a good compromise between completely open ecosystems (nature) and controlled conditions (laboratory) in order to explore expressions that could account for microbial interactions.
